### 2018-07-03 pack/prv/opt/tokipona/aux/article/acm/samplebody-journals.tex
terminada arrumacao minima nos modulos do pacote

\noindent Such procedures are exemplified in \tttt{basic/syntaxhighlighting.py}.

The syntax file has an association of TP
tokens with highlighting groups (HIGs).
It also holds associations of these HIGs to other HIGs 
to use their color settings.
Thus, one may change: associations of tokens to HIGs;
associations of regex patterns to HIGs;
associations of HIGs to other HIGs;
whole color schemes or just sections of it.
The last of these tasks is promptly performed using 
the \tttt{space-c-s} and \tttt{space-c-c} normal Vim commands
supplied by the color plugin of the PRV framework~\cite{vim}.
The other tasks are more easily performed by
scripting Vim, Python import and package source code,
depending on the user ability and intended tasks.

The use of different color schemes within the same syntax
file is exemplified in Figure~\ref{fig:syn}.
By copying only the \tttt{syntax/tokipona.vim} and \tttt{ftdetect/tokipona.vim} files,
a user might benefit from the standard Vim commands.
E.g. \tttt{:colorscheme blue},
\tttt{:colorscheme solarize}, \tttt{:colorscheme gruvbox} or
\tttt{:colorscheme elflord} to see the same text colored
with different color schemes (association of colors to sets of tokens).
To interfere directly on the colors chosen,
\tttt{:highlight Normal guifg=\#00000 guibg=\#0000ff}
will change the standard foreground (text) color to pure black
and background to pure blue.
See \tttt{:h gui-colors} and \tttt{:h highlight}
for the way you might edit colors directly.
One might obtain an HTML file with the colored TP text
using the \tttt{:TOhtml} command.

@@

\subsubsection{Advanced syntax highlighting considerations}
Standard guidelines for syntax highlighting
depend heavily on cultural and use factors
and have scarce scientific studies~\cite{stack}.
There are informed projects such as Solarized~\cite{solarized},
which present solutions for some contexts.
Strikingly, standard guidelines for syntax highlighting were not found.
Therefore, we considered current data visualization
theory~\cite{dv1,dv2,dv3,dv4} to glimpse at the potential guidelines:
\begin{itemize}
  \item The use of blue and other high frequency colors (such as
    violet) for the background.
  \item The avoidance of blue and other high frequency
  colors to fill small objects, such as letters and words.
  \item Explore simplicity and elegance through the use of discrete
  coloring schemes. Most tokens should be preferably of
  the same color.
    One might use only a small number of colors to achieve a clean
    visualization of texts.
  Also, a power-law distribution of tokens among colors might be
    well-suited to mimic natural occurring phenomena.
    (Unfortunately, the TP dictionary classification of words are not in a
    power-law distribution and might be better described by a half-normal or by
overlapping normal distributions over the three peaks: $\sim35-58,
12-18, 1-5$ occurrences).
  \item Physical stimuli might be related to perceptual stimuli both 
    through a power-law or an exponential law (respectively known
    as Steven's law and Weber-Fechner's law).
    This might be useful e.g. for coloring sets of tokens
    considering their similarity.
  \item One usually wishes to maximize contrast,
  although taste and less wearing combinations might
  also dictate the coloring choices.
  \item Stipulate axes of parameters to set colors.
  Wavelength or frequency is an obvious axis given the considerations
    previously exposed, but they are not perceptually uniform.
    Considering perceptually uniform color spaces in making
    choices may be relevant.
  \item One might consider two types of color schemes: those with a dark background,
  which are more comfortable at first; and those with a light background,
  which are usually impressive (and even annoying) at first,
  but the eye gets used to it and it keeps you more stimulated.
\item The blue color is specially related to physiological
  stimulation of the body~\cite{blue,blue2},
    and programmers often report that a blue background keeps them
    awake and more concentrated (I also notice such effect,
    and I have programmed daily for more than 10 years).
  \item Colors have been associated to enhancements in specific tasks,
    e.g. blue and red are respectively associated to enhanced
    creativity and detail-oriented tasks in~\cite{blue}.
  \item It is important to consider that the user might 
    stay many hours
  at a text editor (and we very often do),
    and that the colors and formats involved are
  prone to entail a considerable effect in the body and mental
    activity and thus in the quality of life and work of a writer
    (e.g. a programmer).
  \item Ideally, one should have facilities for tuning 
    the syntax highlighting (e.g. through keyboard shortcuts)
  as envisioned in~\cite{vim}.
\end{itemize}

>>>>

\noindent Such procedures are exemplified in \tttt{basic/syntaxhighlighting.py}.
One may change: associations of tokens to HIGs;
associations of regex patterns to HIGs;
associations of HIGs to other HIGs;
whole color schemes or just sections of it.
The last of these tasks is promptly performed using 
the \tttt{space-c-s} and \tttt{space-c-c} normal Vim commands
supplied by the color plugin of the PRV framework~\cite{vim}.
The other tasks are more easily performed by
scripting Vim, Python import and package source code,
depending on the user ability and intended tasks.
The use of different color schemes within the same syntax
file is exemplified in Figure~\ref{fig:syn}.
By copying only the \tttt{syntax/tokipona.vim} and \tttt{ftdetect/tokipona.vim} files,
a user might benefit from the standard Vim commands.
E.g. \tttt{:colorscheme blue},
\tttt{:colorscheme solarize}, \tttt{:colorscheme gruvbox} or
\tttt{:colorscheme elflord} to see the same text colored
with different color schemes (association of colors to sets of tokens).
To interfere directly on the colors chosen,
\tttt{:highlight Normal guifg=\#00000 guibg=\#0000ff}
will change the standard foreground (text) color to pure black
and background to pure blue.
See \tttt{:h gui-colors} and \tttt{:h highlight}
for the way you might edit colors directly.
One might obtain an HTML file with the colored TP text
using the \tttt{:TOhtml} command.

@@

\subsubsection{Advanced syntax highlighting considerations}
Standard guidelines for syntax highlighting
depend heavily on cultural and use factors
and have scarce scientific studies~\cite{stack}.
There are informed projects such as Solarized~\cite{solarized},
which present solutions for some contexts.
Strikingly, standard guidelines for syntax highlighting were not found.
Therefore, we considered current data visualization
theory~\cite{dv1,dv2,dv3,dv4} to glimpse at the potential guidelines:
\begin{itemize}
  \item The use of blue and other high frequency colors (such as
    violet) for the background and the avoidance of their use
  to fill small objects, such as letters and words.
  \item Explore simplicity and elegance in the  coloring schemes.
    Most tokens should be preferably of the same color, i.e.
    a small number of colors might be explored to achieve a clean
    visualization of texts.
  \item A power-law distribution of tokens among colors might be
    well-suited to mimic natural occurring phenomena.
    The TP dictionary morphosyntactic classification of words
    resemble a long-tailed distribution: as shown in Table~\ref{foobar},
    few classes hold the vast majority of words.
  \item Physical stimuli might be related to perceptual stimuli both 
    through a power-law or an exponential law (respectively known
    as Steven's law and Weber-Fechner's law).
    This might be useful e.g. for coloring sets of tokens
    considering their similarity, or correctly choosing values
    for e.g. the hue or luminosity channels.
  \item One usually wishes to maximize contrast,
  although taste and less wearing combinations might
  also dictate the coloring choices.
  \item Stipulate axes of parameters to set colors.
  Wavelength or frequency is an obvious axis given the considerations
    previously exposed, but they are not perceptually uniform.
    Considering perceptually uniform color spaces in making
    choices may be relevant, such as CIElab and CIEluv.
  \item Consider two types of color schemes: of those with a dark background,
  which are more comfortable at first; and of those with a light background,
  which are usually impressive (and even annoying) at first,
  but the eye adapts in a few minutes, keeps you more stimulated,
    and is suitable for bright contexts (e.g. in daylight) .
\item The blue color is specially related to physiological
  stimulation of the body~\cite{blue,blue2},
    and programmers often report that a blue background keeps them
    awake and more concentrated (the author also notices such effect).
  \item Colors have been associated to enhancements in specific tasks,
    e.g. blue and red are respectively associated to enhanced
    creativity and detail-oriented tasks in~\cite{blue}.
  \item It is important to consider that \emph{a text editor user} might 
    stay many hours using the tool (and they very often do),
    and that the colors and formats involved are
  thus prone to entail a considerable effect in the body and mental
    activity, the quality of life, and work results of a writer
    (e.g. a programmer).
  \item Ideally, one should have facilities for tuning 
    the syntax highlighting (e.g. through keyboard shortcuts)
  as envisioned and implemented in~\cite{vim}.
\end{itemize}
